Figures and results for main manuscript

### 2021-03-18 analysis/rev_prototyping/manuscript_main_rev.Rmd
today

fia_ct <- read.csv(here::here("fia_cts.csv"))

all_ct <- rbind(all_ct, fia_ct)
all_ct <- all_ct %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat))

@@

>>>>

#fia_ct <- read.csv(here::here("fia_cts.csv"))

#all_ct <- rbind(all_ct, fia_ct)
all_ct <- all_ct %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat))

@@

# Very small communities

all_di <- all_di %>%
  mutate( `Number of elements` = ifelse(nparts < 2000, "Less than 2000", "More than 2000"))
diss_results <- all_di %>%
  filter(nparts > 20) %>%
  group_by(Dataset,  `Number of elements` ) %>%
  summarize(high_diss = mean(real_po_percentile_excl > 95),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High dissimilarity` = paste0(signif(100 * high_diss, 2), "%; n = ", nsites)) %>%
  select(-nsites)

skew_results <- all_di %>%
  filter(nparts > 40, s0 > 2) %>%
  group_by(Dataset,  `Number of elements` ) %>%
  summarize(high_skew = mean(skew_percentile_excl > 97.5),
            low_skew = mean(skew_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High skew` = paste0(signif(100 * high_skew, 2), "%; n = ", nsites),
         `Low skew` = paste0(signif(100 * low_skew, 2), "%; n = ", nsites)) %>%
  select(-nsites)

metric_results <- all_di %>%
  filter(nparts > 40) %>%
  group_by(Dataset,  `Number of elements` ) %>%
  summarize(high_simpson = mean(simpson_percentile_excl > 97.5),
            low_simpson = mean(simpson_percentile < 2.5),
            high_shannon = mean(shannon_percentile_excl > 97.5),
            low_shannon = mean(shannon_percentile < 2.5),
            high_nsingletons = mean(nsingletons_percentile_excl > 97.5),
            low_nsingletons = mean(nsingletons_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High Simpson` = paste0(signif(100 * high_simpson, 2), "%; n = ", nsites),
         `Low Simpson` = paste0(signif(100 * low_simpson, 2), "%; n = ", nsites),
         `High Shannon` = paste0(signif(100 * high_shannon, 2), "%; n = ", nsites),
         `Low Shannon` = paste0(signif(100 * low_shannon, 2), "%; n = ", nsites),
         `High proportion of rare species` = paste0(signif(100 * high_nsingletons, 2), "%; n = ", nsites),
         `Low proportion of rare species` = paste0(signif(100 * low_nsingletons, 2), "%; n = ", nsites)) %>%
  select(-nsites)

cols1 <- c("Dataset", "Number of elements", "High dissimilarity", "High proportion of rare species", "High skew", "Low Simpson", "Low Shannon")

plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "sim_pos_from_best", "Dissimilarity") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "real_po_percentile_mean", "Dissimilarity") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "nsingletons_95_ratio_2t", "Number of rare species") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "nsingletons_percentile_mean", "Number of rare species") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "simpson_95_ratio_2t", "Simpson") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "simpson_percentile_mean", "Simpson") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "shannon_95_ratio_2t", "Shannon") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "shannon_percentile_mean", "Shannon") +
  facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

ggplot(all_di, aes(s0, n0, color = `Number of elements`)) +
  geom_point() +
  xlim(0, 30) +
  ylim(0,150)

all_di <- all_di %>%
  mutate( `Number of elements` = ifelse(nparts < 1000, "Less than 1000", "More than 1000"))
diss_results <- all_di %>%
  filter(nparts > 20) %>%
  group_by(Dataset,  `Number of elements`) %>%
  summarize(high_diss = mean(real_po_percentile_excl > 95),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High dissimilarity` = paste0(signif(100 * high_diss, 2), "%; n = ", nsites)) %>%
  select(-nsites)
ggplot(filter(all_di, dat == "mcdb"), aes(skew_95_ratio_2t)) + 
  geom_histogram() +
  facet_wrap(vars(`Number of elements`), scales = "free_y")

skew_results <- all_di %>%
  filter(nparts > 40, s0 > 2) %>%
  group_by(Dataset,  `Number of elements`) %>%
  summarize(high_skew = mean(skew_percentile_excl > 97.5),
            low_skew = mean(skew_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High skew` = paste0(signif(100 * high_skew, 2), "%; n = ", nsites),
         `Low skew` = paste0(signif(100 * low_skew, 2), "%; n = ", nsites)) %>%
  select(-nsites)

metric_results <- all_di %>%
  filter(nparts > 40) %>%
  group_by(Dataset, `Number of elements`) %>%
  summarize(high_simpson = mean(simpson_percentile_excl > 97.5),
            low_simpson = mean(simpson_percentile < 2.5),
            high_shannon = mean(shannon_percentile_excl > 97.5),
            low_shannon = mean(shannon_percentile < 2.5),
            high_nsingletons = mean(nsingletons_percentile_excl > 97.5),
            low_nsingletons = mean(nsingletons_percentile < 2.5),
            nsites = dplyr::n()) %>%
  ungroup() %>%
  mutate(`High Simpson` = paste0(signif(100 * high_simpson, 2), "%; n = ", nsites),
         `Low Simpson` = paste0(signif(100 * low_simpson, 2), "%; n = ", nsites),
         `High Shannon` = paste0(signif(100 * high_shannon, 2), "%; n = ", nsites),
         `Low Shannon` = paste0(signif(100 * low_shannon, 2), "%; n = ", nsites),
         `High proportion of rare species` = paste0(signif(100 * high_nsingletons, 2), "%; n = ", nsites),
         `Low proportion of rare species` = paste0(signif(100 * low_nsingletons, 2), "%; n = ", nsites)) %>%
  select(-nsites)

cols1 <- c("Dataset", "Number of elements", "High dissimilarity", "High proportion of rare species", "High skew", "Low Simpson", "Low Shannon")

# Resampling

all_di_with_resamples <- read.csv(here::here("analysis", "rev_prototyping", "all_di.csv"))

jk_di_full <- read.csv(here::here("analysis", "rev_prototyping", "all_di_jk.csv"))

all_ct_with_resamples <- read.csv(here::here("analysis", "rev_prototyping", "all_ct.csv"))

jk_ct <- read.csv(here::here("analysis", "rev_prototyping", "all_ct_jk.csv"))

jk_di_full <- left_join(jk_di_full, jk_ct)

all_di_with_resamples <- left_join(all_di_with_resamples, all_ct_with_resamples)

jk_di <- jk_di_full %>%
  group_by_all() %>%
  mutate(site_source = unlist(strsplit(site, "_")[[1]][[1]])) %>%
  ungroup() %>%
  select(dat, site, site_source, s0, n0, skew_percentile_excl,skew_percentile, simpson_percentile, simpson_percentile_excl, shannon_percentile, shannon_percentile_excl, nsingletons_percentile, nsingletons_percentile_excl, real_po_percentile, real_po_percentile_excl, sim_pos_from_best, skew_95_ratio_2t, simpson_95_ratio_2t, shannon_95_ratio_2t, nparts)

skewcols <- colnames(jk_di)[ grepl("skew", colnames(jk_di))]

jk_di[ which(jk_di$s0 < 3), skewcols] <- NA

jk_di_means <- jk_di %>%
  filter(n0 != s0,
         s0 != 1,
         n0 != (s0 + 1)) %>%
  group_by(dat, site_source) %>%
  mutate(njks = dplyr::n(),
         njks_skew = sum(!is.na(skew_percentile)))%>%
  summarise_at(vars(-group_cols(), -site), mean, na.rm =  T) %>%
  mutate(resampling = "Subsampling") %>%
  mutate(site = as.numeric(site_source)) %>%
#  select(-site_source) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20,
         njks == 10) %>%
  ungroup()

jk_di[ which(jk_di$njks_skew < 10), skewcols] <- NA

skewcols <- colnames(all_di_with_resamples[ which(grepl("skew", colnames(all_di_with_resamples)))])

all_di_with_resamples[ which(all_di_with_resamples$s0 < 3), skewcols] <- NA

all_di_with_resamples <- all_di_with_resamples %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 != s0,
         s0 != 1,
       #  !singletons,
         n0 != (s0 + 1)) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20) %>%
  mutate(resampling = ifelse(singletons, "Rare species", "Raw")) %>%
  bind_rows(jk_di_means)

metric_rs_results <- all_di_with_resamples %>%
  group_by(Dataset, resampling) %>%
  summarize(
    skew_high = mean(skew_percentile_excl > 97.5, na.rm = T),
    skew_low = mean(skew_percentile < 2.5, na.rm = T),
    even_high = mean(simpson_percentile_excl > 97.5, na.rm = T),
    even_low = mean(simpson_percentile < 2.5, na.rm = T),
    shannon_high = mean(shannon_percentile_excl > 97.5, na.rm = T),
    shannon_low = mean(shannon_percentile < 2.5, na.rm = T),
    nsingletons_high = mean(nsingletons_percentile_excl > 97.5, na.rm = T),
    nsingletons_low = mean(nsingletons_percentile < 2.5, na.rm = T),
    nskew = sum(!is.na(skew_percentile)),
    nother = dplyr::n()
  ) %>%
  ungroup()

diss_rs_results <- all_di_with_resamples %>%
  filter(nparts > 40) %>%
  group_by(Dataset, resampling) %>%
  summarize(dissimilarity_high = mean(real_po_percentile_excl > 95, na.rm = T),
            ndiss = sum(!is.na(real_po_percentile_excl))) %>%
  ungroup()

rs_results <- left_join(metric_rs_results, diss_rs_results)

#metric_cols <- colnames(rs_results [ grepl("_", colnames(rs_results))])

as.perc <- function(x) {
  return((signif(100 * x, digits = 2)))
}

rs_results_desc <- rs_results %>%
  mutate(`High dissimilarity` = paste0(as.perc(dissimilarity_high), "%; n =", ndiss),
         `High proportion of rare species` = paste0(as.perc(nsingletons_high), "%; n =", nother),
         `High skew` = paste0(as.perc(skew_high), "%; n = ", nskew),
         `Low Simpson` = paste0(as.perc(even_low), "%; n = ", nother),
         `Low Shannon` = paste0(as.perc(shannon_low), "%; n = ", nother),
         `Low proportion of rare species` = paste0(as.perc(nsingletons_low), "%; n =", nother),
         `Low skew` = paste0(as.perc(skew_low), "%; n = ", nskew),
         `High Simpson` = paste0(as.perc(even_high), "%; n = ", nother),
         `High Shannon` = paste0(as.perc(shannon_high), "%; n = ", nother),
         `Resampling scheme` = ordered(resampling, levels = c("Raw", "Subsampling", "Rare species"))) %>%
  arrange(Dataset, `Resampling scheme`)

cols1 <- c("Dataset", "Resampling scheme", "High dissimilarity", "High proportion of rare species", "High skew", "Low Simpson", "Low Shannon")

cols2 <- c("Dataset","Resampling scheme",  "Low proportion of rare species", "Low skew", "High Simpson", "High Shannon")

rs_results_desc %>%
  select(cols1)

rs_results_desc %>%
  select(cols2)
```
```{r, fig.dim = c(6,4)}
rs_results_long <- rs_results %>% 
  tidyr::pivot_longer((-c("Dataset","resampling")), names_to = "metric", values_to = "value") %>%
  filter(grepl("_", metric)) %>%
  mutate(ometric = ordered(metric, levels = c("skew_high", "skew_low", "even_low", "even_high", "shannon_low", "shannon_high", "nsingletons_high", "nsingletons_low", "dissimilarity_high")),
         ometric2 = ordered(metric, levels = c("dissimilarity_high",  "nsingletons_high", "skew_high",  "even_low",  "shannon_low",  "nsingletons_low","skew_low","even_high","shannon_high"))) %>%
  mutate(weird_dir = as.numeric(ometric2) > 5,
         percentile_cutoff = ifelse(grepl("diss", metric), .05, .025),`Resampling scheme` = ordered(resampling, levels = c("Raw", "Subsampling", "Rare species")))

metric_names <- rs_results_long %>%
  select(ometric2) %>%
  distinct() %>%
  arrange(ometric2) %>%
  mutate(metric_desc =(c("High dissimilarity", "High proportion \nof rare species", "High skew", "Low Simpson", "Low Shannon", "Low proportion \nof rare species", "Low skew", "High Simpson", "High Shannon"))) %>%
  mutate(metric_desc = ordered(metric_desc))

rs_results_long <- left_join(rs_results_long, metric_names)
# 
# ggplot(filter(rs_results_long, !weird_dir), aes(metric_desc, value, color = `Resampling scheme`)) +
#   geom_point() +
#   facet_wrap(vars(Dataset), scales = "fixed", ncol = 1) +
#   geom_point(aes(y = percentile_cutoff), shape = 95, color = "black", size = 10) +
#   scale_color_viridis_d(end = .8) +
#   ylab("Proportion of extreme observed values") +
#   xlab("")

gridExtra::grid.arrange(grobs = list(
  ggplot(filter(rs_results_long, !grepl("Misc", Dataset), !weird_dir), aes(metric_desc, value, color = `Resampling scheme`)) +
  geom_point() +
  facet_wrap(vars(Dataset), scales = "fixed", ncol =2) +
  geom_point(aes(y = percentile_cutoff), shape = 95, color = "black", size = 10) +
  scale_color_viridis_d(end = .8) +
  ylab("Proportion of extreme observed values") +ylim(0, .7) +
  xlab("") + theme(legend.position = "none", axis.text.x = element_text(angle = 90)) ,
  ggplot(filter(rs_results_long, grepl("Misc", Dataset), !weird_dir), aes(metric_desc, value, color = `Resampling scheme`)) +
  geom_point() +
  facet_wrap(vars(Dataset), scales = "fixed", ncol = 2) +
  geom_point(aes(y = percentile_cutoff), shape = 95, color = "black", size = 10) +
  scale_color_viridis_d(end = .8) +
  ylab("") +
  xlab("") + ylim(0, .7) +
  theme(legend.position = "bottom", legend.direction = "vertical" , legend.margin = margin(b = .3, r = 0, t = 0, l = 0, unit = "in"), axis.text.x = element_text(angle = 90))), ncol = 2, widths = c( 4,2), bottom = textGrob("Metric"))

gridExtra::grid.arrange(grobs = list(
  ggplot(filter(rs_results_long, !grepl("Misc", Dataset), weird_dir), aes(metric_desc, value, color = `Resampling scheme`)) +
  geom_point() +
  facet_wrap(vars(Dataset), scales = "fixed", ncol =2) +
  geom_point(aes(y = percentile_cutoff), shape = 95, color = "black", size = 10) +
  scale_color_viridis_d(end = .8) +
  ylab("Proportion of extreme observed values") +ylim(0, .7) +
  xlab("") + theme(legend.position = "none", axis.text.x = element_text(angle = 90)) ,
  ggplot(filter(rs_results_long, grepl("Misc", Dataset), weird_dir), aes(metric_desc, value, color = `Resampling scheme`)) +
  geom_point() +
  facet_wrap(vars(Dataset), scales = "fixed", ncol = 2) +
  geom_point(aes(y = percentile_cutoff), shape = 95, color = "black", size = 10) +
  scale_color_viridis_d(end = .8) +
  ylab("") +
  xlab("") + ylim(0, .7) +
  theme(legend.position = "bottom", legend.direction = "vertical" , legend.margin = margin(b = .3, r = 0, t = 0, l = 0, unit = "in"), axis.text.x = element_text(angle = 90))), ncol = 2, widths = c( 4,2), bottom = textGrob("Metric"))

all_di_with_resamples <- all_di_with_resamples %>%
  group_by_all() %>%
  mutate(real_po_percentile_mean = mean(real_po_percentile, real_po_percentile_excl, na.rm = T),
         skew_percentile_mean = mean(skew_percentile, skew_percentile_excl, na.rm = T),
         simpson_percentile_mean = mean(simpson_percentile,simpson_percentile_excl, na.rm = T),
         shannon_percentile_mean = mean(shannon_percentile, shannon_percentile_excl, na.rm = T),
         nsingletons_percentile_mean = mean(nsingletons_percentile, nsingletons_percentile_excl)) %>%
  ungroup()

  plot_percentile_hist(filter(all_di_with_resamples, resampling == "Subsampling"), "real_po_percentile_mean", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset", tails = 1),
  plot_percentile_hist(filter(all_di_with_resamples, resampling == "Subsampling"), "nsingletons_percentile_mean", "\nNumber of rare species", facetvar = "Dataset"),
  plot_percentile_hist(filter(all_di_with_resamples, resampling == "Subsampling"), "skew_percentile_mean", "\nSkewness", facetvar = "Dataset", min_s0 = 3),
  plot_percentile_hist(filter(all_di_with_resamples, resampling == "Subsampling"), "simpson_percentile_mean", "\nSimpson evenness", facetvar = "Dataset"),
  plot_percentile_hist(filter(all_di_with_resamples, resampling == "Subsampling"), "shannon_percentile_mean", "\nShannon diversity", facetvar = "Dataset")
), ncol = 5,
top = textGrob("Subsampling results", gp = gpar(fill = "white")), left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")),
bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white"))

  plot_percentile_hist(filter(all_di_with_resamples, resampling == "Rare species"), "real_po_percentile_mean", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset", tails = 1),
  plot_percentile_hist(filter(all_di_with_resamples, resampling == "Rare species"), "nsingletons_percentile_mean", "\nNumber of rare species", facetvar = "Dataset"),
  plot_percentile_hist(filter(all_di_with_resamples, resampling == "Rare species"), "skew_percentile_mean", "\nSkewness", facetvar = "Dataset", min_s0 = 3),
  plot_percentile_hist(filter(all_di_with_resamples, resampling == "Rare species"), "simpson_percentile_mean", "\nSimpson evenness", facetvar = "Dataset"),
  plot_percentile_hist(filter(all_di_with_resamples, resampling == "Rare species"), "shannon_percentile_mean", "\nShannon diversity", facetvar = "Dataset")
), ncol = 5,
top = textGrob("Adjusted for rare species results", gp = gpar(fill = "white")), left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")),
bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white"))
```
